---
title: "Amphiprion percula Larval Trait Heritability"
author: "E Schlatter"
date: "12/18/2021"
output: html_document
---

```{r}
library(tidyverse)
library(MasterBayes) #for pedigree functions
library(MCMCglmm)
library(MCMCpack) #for inverse wishart distributions
library(extraDistr) #for inverse gamma distribution (plotting priors)
library(bayesplot) #mcmc_areas plot function
```


##Read in and process the data.

Creates the following objects we'll use going forward:

$\textbf{phens}$: a list of each larva and its swimming speed and clutch, plus other covariates to be added. (Each parent also has a row in this list, to be compatible with relatedness matrix, but its swimming speed and covariates are all NA.)

$\textbf{Ainv}$: the inverse of the relatedness matrix, A. In A, $A_{ij}=$probability of identity by descent of an allele in individuals i and j.
```{r data cleanup}
#swim trial data
larvae <- as.data.frame(read.csv('data/Larvae.csv',header=TRUE)) %>%
  mutate(larvalID=Clutch+.01*Larva) #create unique identifier for each larva

#clutch data
metadata <- as.data.frame(read.csv('data/metadata.csv',header=TRUE))
colnames(metadata)[1]='Clutch'

#length data
lengths <- as.data.frame(read.csv('data/Lengths.csv',header=TRUE)) %>%
  mutate(larvalID=Clutch+.01*Larva) %>% #create unique identifier for each larva
  dplyr::select(-Clutch,-Larva) %>%
  rename(Measurer=Initials) %>%
  rename(SL=Average.length) %>%
  mutate(SL=as.numeric(SL))  #this will give "NAs introduced by coercion" -- that's fine

for(i in 1:nrow(larvae)){ #remove SL measurements for dead larvae; they're not accurate
  larvae$SL[i]=ifelse(larvae$Dead.before.preservation.[i]==1,NA,larvae$SL[i])} 

#merge swim trial and length data
larvae <- merge(larvae,lengths,by='larvalID')

larvae <- larvae %>% mutate('Time'=Time.min.+(1/60)*Time.sec.) %>% #add a Time column
  subset(Clutch>4) %>% #first four clutches were practice, and we won't use them in the dataset
  mutate(Clutch=as.factor(Clutch)) %>%
  mutate(larvalID=as.factor(larvalID))

#merge the larvae and metadata data frames, so metadata info is associated with each larva
larvae <- merge(larvae,metadata,by='Clutch') %>%
  dplyr::select(larvalID,everything()) #move larvalID to first column

larvae <- larvae %>% subset(is.na(SL)==FALSE) %>% #remove larvae with no SL measurement
  subset(is.na(Time)==FALSE) %>% #remove larvae with no speed measurement
  mutate(speed=Time-2) #Ucrit is swim time minus 2cm/sec

#make sure all covariates are in the right format
larvae <- larvae %>%
  mutate(R1.ID.=as.factor(R1.ID.)) %>%
  mutate(R2.ID.=as.factor(R2.ID.)) %>%
  mutate(Date=as.Date(Date, "%d-%b-%y")) %>%
  mutate(PostSwitch=as.factor(PostSwitch)) %>%
  mutate(Parental.tank=as.factor(Parental.tank))
```
Trim the dataset to larvae with both maternal and paternal half-sibs. 581 larvae from 12 mothers and 12 fathers.
```{r trim data}
mothers_trim <- c(5,6,9,10,13,18,22,30,59,66,70,114) #mothers in the trimmed dataset
#error in original data entry: father #137 is sometimes written #131
fathers_trim <- c(38,96,97,98,103,108,111,115,137,156,165,171) #fathers in the trimmed dataset
metadata <- subset(metadata,R1.ID.%in%mothers_trim & R2.ID.%in%fathers_trim)
larvae <- subset(larvae,R1.ID.%in%mothers_trim & R2.ID.%in%fathers_trim)
```

Create:
  - pedigree (and full_pedig, including the parents)
  - Ainv
  - phens (animal, speed, length, covariates)

```{r}
#pedigree and Ainv -- we'll need this later for heritability analysis
pedigree <- dplyr::select(larvae,id=larvalID,dam=R1.ID.,sire=R2.ID.)

#pedigree utility functions to insert parents and correctly order, from MasterBayes.
full_pedig <- pedigree %>%
  insertPed() %>%
  orderPed()
Ainv <- MCMCglmm::inverseA(full_pedig)$Ainv

#phenotype data frame
phens <- dplyr::select(larvae,animal=larvalID,speed=speed, length=SL, clutch=Clutch, 
                       dam=R1.ID., sire=R2.ID., tank=Parental.tank, age=Age.at.hatch)
  #add empty rows to phens for parents
parents <- subset(full_pedig,is.na(dam)) %>%
  transmute(animal=id,speed=as.numeric(dam),length=as.numeric(dam),clutch=NA,
            dam=NA,sire=NA,tank=NA,age=1)
phens <- as.data.frame(phens)
phens <- rbind(parents,phens)
```

## Animal model

### Priors
From Hadfield course notes (p.13):
"The inverse gamma is a special case of the inverse Wishart, although it is parametrised using shape and scale, where nu = 2$*$shape and V = scale/shape (or shape = nu/2 and scale = nu$*$V/2 ). MCMCpack provides a density function (dinvgamma) for the inverse gamma distribution."

"The distribution tends to a point mass on V as the degree of belief parameter, nu, goes to infinity." The mean is not define for nu<2.

Several priors to try:
```{r priors, results='hide'}

#parameter-expanded prior
prior_expand_params <- list(V = 1,nu = 1000, alpha.mu=0, alpha.V=1)

#also parameter-expanded, from tuto_en; with residual edited (nu=0.002 instead of nu=1) to be proper
prior_ext_params <- list(V=1,nu=1,alpha.mu=0,alpha.V=1000)

#inverse gamma prior
prior_invgamma_params <- list(V = 1,nu = .002)

#gentler inverse gamma prior
prior_invgamma2_params <- list(V = 1,nu = .02)

#flat prior
prior_flat_params <- list(V = 1e-10,nu = -1) #trusting based on course notes that this is actually flat, because I can't tell what a negative nu means
```


Plot them (when possible):
```{r}
#invgamma
param_list <- prior_invgamma_params
nu <- param_list$nu
V <- param_list$V
#corresponding shape and scale (invgamma) parameters
shape=nu/2
scale=nu*V/2

#invgamma2
param_list.2 <- prior_invgamma2_params
nu.2 <- param_list.2$nu
V.2 <- param_list.2$V
#corresponding shape and scale (invgamma) parameters
shape.2=nu.2/2
scale.2=nu.2*V.2/2

xrange <- seq(0.001,1,.001)

ggplot()+
  geom_line(aes(x=xrange,y=MCMCpack::dinvgamma(xrange,shape,scale),
                color='invgamma'))+
  geom_line(aes(x=xrange,y=MCMCpack::dinvgamma(xrange,shape.2,scale.2),
                color='invgamma2'))+
  labs(x="sigma^2",y="probability",title='prior distributions')
```


### SL:animal model
Model: y=a+c+e  (a=additive genetic effect; c=clutch; e=residual)

```{r}
param_list=prior_invgamma2_params
model_statement="animal+clutch"
n_vars=2 #not including the residual (units)
n_mcmc_iter=100000


model_run <- function(param_list,model_statement,n_vars,n_mcmc_iter){
  #structure for plugging in the params to generate a full prior -- this is what MCMCglmm takes
  G <- list()
  for(i in 1:n_vars){
    varname <- paste("G",i,sep="")
    G[[varname]] <- param_list
    }
  prior <- list(R=param_list, #variance prior for residual
                G=G)  #variance priors for random effects

  #run the model
  model <- MCMCglmm(length ~ 1+age,
                        random=formula(paste0('~',model_statement)),
                        prior=prior,
                        data=phens,nitt=n_mcmc_iter,burnin=round(.01*n_mcmc_iter),thin=10,verbose=TRUE)
          #I think the "missing values in random predictors" warning is just because the parents in phens don't have predictor values
  return(model)
}

#function that takes an mcmc model object (output of MCMCglmm) and looks at the results
model_diagnostics <- function(model){
  #trace and density plots
  plot(model)
  
  #store convergence info by variable (original and scaled)
  diagnostics=data.frame()
    #original variables
  for(i in 1:ncol(model$VCV)){
    v_name <- colnames(model$VCV)[i]
    v_mcmc <- model$VCV[,i]
    v_diagnostics <- data.frame(variable=v_name,
                                effSamp=effectiveSize(v_mcmc),
                                firstAutocorr=autocorr(v_mcmc)[2],lastAutocorr=autocorr(v_mcmc)[nrow(autocorr(v_mcmc))])
    diagnostics=rbind(diagnostics,v_diagnostics)
  }
    #scaled variables  
  for(i in 1:ncol(matrix_scaled_vars)){
    v_name <- colnames(matrix_scaled_vars)[i]
    v_mcmc <- scaled_vars[[i]]
    v_diagnostics <- data.frame(variable=v_name,
                     effSamp=effectiveSize(v_mcmc),
                     firstAutocorr=autocorr(v_mcmc)[2],lastAutocorr=autocorr(v_mcmc)[nrow(autocorr(v_mcmc))])
    diagnostics=rbind(diagnostics,v_diagnostics)
  }
  rownames(diagnostics)=c(colnames(model$VCV),colnames(matrix_scaled_vars))
  
  return(diagnostics)
}

model_results <- function(model){
  #proportional variances as new mcmc objects
  scaled_vars <- list()
  for(i in 1:ncol(model$VCV)){
    varname <- paste('scale',colnames(model$VCV)[i],sep='_')
    scaled_vars[[varname]] <- model$VCV[,i]/rowSums(as.data.frame(model$VCV)) #store them in a list
  }
    
  #plot histograms of variance estimates
  matrix_scaled_vars <- do.call(cbind,scaled_vars) #put all variables together into a matrix for plotting
  plot(mcmc_areas(matrix_scaled_vars,prob=0.95,area_method='scaled height')) #scaled to 1
  
  #store estimates (mean, median, upper+lower95) for each variable (original and scaled)
  estimates=data.frame()
    #original variables
  for(i in 1:ncol(model$VCV)){
    v_name <- colnames(model$VCV)[i]
    v_mcmc <- model$VCV[,i]
    v_estimates <- data.frame(variable=v_name,
                            mean=mean(v_mcmc),median=median(v_mcmc),upper95=HPDinterval(v_mcmc)[,2],lower95=HPDinterval(v_mcmc)[,1])
    estimates=rbind(estimates,v_estimates)
  }
    #scaled variables  
  for(i in 1:ncol(matrix_scaled_vars)){
    v_name <- colnames(matrix_scaled_vars)[i]
    v_mcmc <- scaled_vars[[i]]
    v_estimates <- data.frame(variable=v_name,
                            mean=mean(v_mcmc),median=median(v_mcmc),upper95=HPDinterval(v_mcmc)[,2],lower95=HPDinterval(v_mcmc)[,1])
    estimates=rbind(estimates,v_estimates)
  }
  rownames(estimates)=c(colnames(model$VCV),colnames(matrix_scaled_vars))
  
  #estimates_long <- pivot_longer(estimates,mean:lower95,names_to='SummStat',values_to='Value') #in "long" format -- not sure yet which will be most useful

 return(estimates)

}

model <- model_run(prior_invgamma2_params,"animal+clutch",2,1000)
res <- model_results(a)
```

