---
title: "Process larval trait data"
author: "E Schlatter"
date: "1/14/2022"
output: html_document
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(MasterBayes) #for pedigree functions
library(MCMCglmm) #for running the model
library(MCMCpack) #for inverse wishart distributions
library(bayesplot) #mcmc_areas plot function
library(gridExtra) #multiple plots per figure
```

## Read in and process the data.

Creates the following objects we'll use going forward:

$\textbf{phens}$: a list of each larva and its swimming speed and clutch, plus other covariates to be added. (Each parent also has a row in this list, to be compatible with relatedness matrix, but its swimming speed and covariates are all NA.)

$\textbf{Ainv}$: the inverse of the relatedness matrix, A. In A, $A_{ij}=$probability of identity by descent of an allele in individuals i and j.
```{r data cleanup}
#swim trial data
larvae <- as.data.frame(read.csv('../data/Larvae.csv',header=TRUE)) %>%
  mutate(larvalID=Clutch+.01*Larva) #create unique identifier for each larva
##start with 1567 larvae

#clutch data
metadata <- as.data.frame(read.csv('../data/metadata.csv',header=TRUE))
colnames(metadata)[1]='Clutch'

#length data
lengths <- as.data.frame(read.csv('../data/Lengths.csv',header=TRUE)) %>%
  mutate(larvalID=Clutch+.01*Larva) %>% #create unique identifier for each larva
  dplyr::select(-Clutch,-Larva) %>%
  rename(Measurer=Initials) %>%
  rename(SL=Average.length) %>%
  mutate(SL=as.numeric(SL))  #this will give "NAs introduced by coercion" -- that's fine

#merge swim trial and length data
larvae <- merge(larvae,lengths,by='larvalID',all.x=TRUE)

for(i in 1:nrow(larvae)){ #remove SL measurements for dead larvae; they're not accurate
  larvae$SL[i]=ifelse(larvae$Dead.before.preservation.[i]=="",larvae$SL[i],NA)}

larvae <- larvae %>% mutate('Time'=Time.min.+(1/60)*Time.sec.) %>% #add a Time column
  subset(Clutch>4) %>% #first four clutches were practice, and we won't use them in the dataset
  mutate(Clutch=as.factor(Clutch)) %>%
  mutate(larvalID=as.factor(larvalID))
##removed first four clutches (120 larvae) --> 1447 larvae left

#merge the larvae and metadata data frames, so metadata info is associated with each larva
larvae <- merge(larvae,metadata,by='Clutch') %>%
  dplyr::select(larvalID,everything()) #move larvalID to first column

larvae <- larvae %>% subset(is.na(SL)==FALSE) %>% #remove larvae with no SL measurement
  mutate(speed=Time-2) #Ucrit is swim time minus 2cm/sec
##removed larvae with no SL (134 larvae) --> 1313 larvae left

#make sure all covariates are in the right format
larvae <- larvae %>%
  mutate(R1.ID.=as.factor(R1.ID.)) %>%
  mutate(R2.ID.=as.factor(R2.ID.)) %>%
  mutate(Date=as.Date(Date, "%d-%b-%y")) %>%
  mutate(PostSwitch=as.factor(PostSwitch)) %>%
  mutate(Parental.tank=as.factor(Parental.tank))
```
Trim the dataset to larvae with both maternal and paternal half-sibs. 629 larvae from 12 mothers and 12 fathers.
```{r trim data}
mothers_trim <- c(5,6,9,10,13,18,22,30,59,66,70,114) #mothers in the trimmed dataset
fathers_trim <- c(38,96,97,98,103,108,111,115,137,156,165,171) #fathers in the trimmed dataset
  #error in original data entry: father #137 is sometimes written #131
metadata <- subset(metadata,R1.ID.%in%mothers_trim & R2.ID.%in%fathers_trim)
larvae <- subset(larvae,R1.ID.%in%mothers_trim & R2.ID.%in%fathers_trim)

##trimmed dataset has 629 larvae
```

Create:

  - pedigree (and full_pedig, including the parents)

  - Ainv

  - phens (animal, speed, length, covariates)

```{r}
#pedigree and Ainv -- we'll need this later for heritability analysis
pedigree <- dplyr::select(larvae,id=larvalID,dam=R1.ID.,sire=R2.ID.)

#pedigree utility functions to insert parents and correctly order, from MasterBayes.
full_pedig <- pedigree %>%
  insertPed() %>%
  orderPed()
Ainv <- MCMCglmm::inverseA(full_pedig)$Ainv

#phenotype data frame
phens <- dplyr::select(larvae,animal=larvalID,speed=speed, length=SL, clutch=Clutch, 
                       dam=R1.ID., sire=R2.ID., tank=Parental.tank, age=Age.at.hatch)
  #add empty rows to phens for parents
parents <- subset(full_pedig,is.na(dam)) %>%
  transmute(animal=id,speed=as.numeric(dam),length=as.numeric(dam),clutch=NA,
            dam=NA,sire=NA,tank=NA,age=1)
phens <- as.data.frame(phens)
phens <- rbind(parents,phens)

#larvae_out <- larvae %>% dplyr::select(larvalID,speed,SL,dam=R1.ID.,sire=R2.ID.,tank=Parental.tank,PostSwitch,age=Age.at.hatch)
#write.csv(larvae_out,file='../data/larvae_all.csv')
#write.csv(phens,file='../data/phens_notrim.csv')
#save(phens,Ainv,list=c('phens','Ainv'),file='../data/h2data.RData')
```
