---
title: "Process data and run models"
author: "E Schlatter"
date: "12/18/2021"
output: html_document
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(MasterBayes) #for pedigree functions
library(MCMCglmm) #for running the model
library(MCMCpack) #for inverse wishart distributions
library(bayesplot) #mcmc_areas plot function
```

```{r}
source("../functions/model_fns.R")
source("priors.R")
```

## Initial model runs
One chain each for all combinations of:
  - priors (5)
  - parental effects or not (2)
  - trimmed/untrimmed data (2)
  - response variable SL or Ucrit (2)

```{r, eval=FALSE}
##test to make sure it works
param_list=prior_invgamma2_params
model_statement="animal+clutch"
n_vars=2 #not including the residual (units)
n_mcmc_iter=100000


model <- model_run(prior_invgamma2_params,'speed',"animal+clutch",2,1000)
save(model,file='../model_runs/test.RData')
diagn <- model_diagnostics(model)
results <- model_results(model)
```

### Full dataset models
```{r, eval=FALSE}
### SL, no parental effects
  #prior_expand
  SL_noparent_fulldata_priorexpand_10e5 <- model_run(prior_expand_params,'length',"animal+clutch",2,100000)
  #prior_ext
  SL_noparent_fulldata_priorext_10e5 <- model_run(prior_ext_params,'length',"animal+clutch",2,100000)
  #invgamma
  SL_noparent_fulldata_invgamma_10e5 <- model_run(prior_invgamma_params,'length',"animal+clutch",2,100000)
  #invgamma2
  SL_noparent_fulldata_invgamma2_10e5 <- model_run(prior_invgamma2_params,'length',"animal+clutch",2,100000)
  #flat
  SL_noparent_fulldata_flatprior_10e5 <- model_run(prior_flat_params,'length',"animal+clutch",2,100000)
  #save them as a list
  SL_noparent_fulldata_10e5 <- list(SL_noparent_fulldata_priorexpand_10e5,SL_noparent_fulldata_priorext_10e5,
                                    SL_noparent_fulldata_invgamma_10e5,SL_noparent_fulldata_invgamma2_10e5,
                                    SL_noparent_fulldata_flatprior_10e5)
  save(SL_noparent_fulldata_10e5,file='../model_runs/SL_noparent_fulldata_10e5.RData')

### SL, with parental effects
  #prior_expand
  SL_parent_fulldata_priorexpand_10e5 <- model_run(prior_expand_params,'length',"animal+clutch+dam+sire",4,100000)
  #prior_ext
  SL_parent_fulldata_priorext_10e5 <- model_run(prior_ext_params,'length',"animal+clutch+dam+sire",4,100000)
  #invgamma
  SL_parent_fulldata_invgamma_10e5 <- model_run(prior_invgamma_params,'length',"animal+clutch+dam+sire",4,100000)
  #invgamma2
  SL_parent_fulldata_invgamma2_10e5 <- model_run(prior_invgamma2_params,'length',"animal+clutch+dam+sire",4,100000)
  #flat -- doesn't work; remove it
  # SL_parent_fulldata_flatprior_10e5 <- model_run(prior_flat_params,'length',"animal+clutch+dam+sire",4,100000)
  #save them as a list
  SL_parent_fulldata_10e5 <- list(SL_parent_fulldata_priorexpand_10e5,SL_parent_fulldata_priorext_10e5,
                                    SL_parent_fulldata_invgamma_10e5,SL_parent_fulldata_invgamma2_10e5)
  save(SL_parent_fulldata_10e5,file='../model_runs/SL_parent_fulldata_10e5.RData')

### Ucrit, no parental effects
  #prior_expand
  Ucrit_noparent_fulldata_priorexpand_10e5 <- model_run(prior_expand_params,'speed',"animal+clutch",2,100000)
  #prior_ext
  Ucrit_noparent_fulldata_priorext_10e5 <- model_run(prior_ext_params,'speed',"animal+clutch",2,100000)
  #invgamma
  Ucrit_noparent_fulldata_invgamma_10e5 <- model_run(prior_invgamma_params,'speed',"animal+clutch",2,100000)
  #invgamma2
  Ucrit_noparent_fulldata_invgamma2_10e5 <- model_run(prior_invgamma2_params,'speed',"animal+clutch",2,100000)
  #flat
  Ucrit_noparent_fulldata_flatprior_10e5 <- model_run(prior_flat_params,'speed',"animal+clutch",2,100000)
  #save them as a list
  Ucrit_noparent_fulldata_10e5 <- list(Ucrit_noparent_fulldata_priorexpand_10e5,Ucrit_noparent_fulldata_priorext_10e5,
                                    Ucrit_noparent_fulldata_invgamma_10e5,Ucrit_noparent_fulldata_invgamma2_10e5,
                                    Ucrit_noparent_fulldata_flatprior_10e5)
  save(Ucrit_noparent_fulldata_10e5,file='../model_runs/Ucrit_noparent_fulldata_10e5.RData')

### Ucrit, with parental effects
  #prior_expand
  Ucrit_parent_fulldata_priorexpand_10e5 <- model_run(prior_expand_params,'speed',"animal+clutch+dam+sire",4,100000)
  #prior_ext
  Ucrit_parent_fulldata_priorext_10e5 <- model_run(prior_ext_params,'speed',"animal+clutch+dam+sire",4,100000)
  #invgamma
  Ucrit_parent_fulldata_invgamma_10e5 <- model_run(prior_invgamma_params,'speed',"animal+clutch+dam+sire",4,100000)
  #invgamma2
  Ucrit_parent_fulldata_invgamma2_10e5 <- model_run(prior_invgamma2_params,'speed',"animal+clutch+dam+sire",4,100000)
  #flat
  Ucrit_parent_fulldata_flatprior_10e5 <- model_run(prior_flat_params,'speed',"animal+clutch+dam+sire",4,100000)
  #save them as a list
  Ucrit_parent_fulldata_10e5 <- list(Ucrit_parent_fulldata_priorexpand_10e5,Ucrit_parent_fulldata_priorext_10e5,
                                    Ucrit_parent_fulldata_invgamma_10e5,Ucrit_parent_fulldata_invgamma2_10e5,
                                    Ucrit_parent_fulldata_flatprior_10e5)
  save(Ucrit_parent_fulldata_10e5,file='../model_runs/Ucrit_parent_fulldata_10e5.RData')
```

### Trimmed dataset models
```{r, eval=FALSE}
### SL, no parental effects
  #prior_expand
  SL_noparent_trimdata_priorexpand_10e5 <- model_run(prior_expand_params,'length',"animal+clutch",2,100000)
  #prior_ext
  SL_noparent_trimdata_priorext_10e5 <- model_run(prior_ext_params,'length',"animal+clutch",2,100000)
  #invgamma
  SL_noparent_trimdata_invgamma_10e5 <- model_run(prior_invgamma_params,'length',"animal+clutch",2,100000)
  #invgamma2
  SL_noparent_trimdata_invgamma2_10e5 <- model_run(prior_invgamma2_params,'length',"animal+clutch",2,100000)
  #flat
  SL_noparent_trimdata_flatprior_10e5 <- model_run(prior_flat_params,'length',"animal+clutch",2,100000)
  #save them as a list
  SL_noparent_trimdata_10e5 <- list(SL_noparent_trimdata_priorexpand_10e5,SL_noparent_trimdata_priorext_10e5,
                                    SL_noparent_trimdata_invgamma_10e5,SL_noparent_trimdata_invgamma2_10e5,
                                    SL_noparent_trimdata_flatprior_10e5)
  save(SL_noparent_trimdata_10e5,file='../model_runs/SL_noparent_trimdata_10e5.RData')

### SL, with parental effects
  #prior_expand
  SL_parent_trimdata_priorexpand_10e5 <- model_run(prior_expand_params,'length',"animal+clutch+dam+sire",4,100000)
  #prior_ext
  SL_parent_trimdata_priorext_10e5 <- model_run(prior_ext_params,'length',"animal+clutch+dam+sire",4,100000)
  #invgamma
  SL_parent_trimdata_invgamma_10e5 <- model_run(prior_invgamma_params,'length',"animal+clutch+dam+sire",4,100000)
  #invgamma2
  SL_parent_trimdata_invgamma2_10e5 <- model_run(prior_invgamma2_params,'length',"animal+clutch+dam+sire",4,100000)
  #flat -- Error: "Mixed model equations singular: use a (stronger) prior"
  #SL_parent_trimdata_flatprior_10e5 <- model_run(prior_flat_params,'length',"animal+clutch+dam+sire",4,100000)
  #save them as a list
  SL_parent_trimdata_10e5 <- list(SL_parent_trimdata_priorexpand_10e5,SL_parent_trimdata_priorext_10e5,
                                    SL_parent_trimdata_invgamma_10e5,SL_parent_trimdata_invgamma2_10e5)
  save(SL_parent_trimdata_10e5,file='../model_runs/SL_parent_trimdata_10e5.RData')


### Ucrit, no parental effects
  #prior_expand
  Ucrit_noparent_trimdata_priorexpand_10e5 <- model_run(prior_expand_params,'speed',"animal+clutch",2,100000)
  #prior_ext
  Ucrit_noparent_trimdata_priorext_10e5 <- model_run(prior_ext_params,'speed',"animal+clutch",2,100000)
  #invgamma
  Ucrit_noparent_trimdata_invgamma_10e5 <- model_run(prior_invgamma_params,'speed',"animal+clutch",2,100000)
  #invgamma2
  Ucrit_noparent_trimdata_invgamma2_10e5 <- model_run(prior_invgamma2_params,'speed',"animal+clutch",2,100000)
  #flat
  Ucrit_noparent_trimdata_flatprior_10e5 <- model_run(prior_flat_params,'speed',"animal+clutch",2,100000)
  #save them as a list
  Ucrit_noparent_trimdata_10e5 <- list(Ucrit_noparent_trimdata_priorexpand_10e5,Ucrit_noparent_trimdata_priorext_10e5,
                                    Ucrit_noparent_trimdata_invgamma_10e5,Ucrit_noparent_trimdata_invgamma2_10e5,
                                    Ucrit_noparent_trimdata_flatprior_10e5)
  save(Ucrit_noparent_trimdata_10e5,file='../model_runs/Ucrit_noparent_trimdata_10e5.RData')

### Ucrit, with parental effects
  #prior_expand
  Ucrit_parent_trimdata_priorexpand_10e5 <- model_run(prior_expand_params,'speed',"animal+clutch+dam+sire",4,100000)
  #prior_ext
  Ucrit_parent_trimdata_priorext_10e5 <- model_run(prior_ext_params,'speed',"animal+clutch+dam+sire",4,100000)
  #invgamma
  Ucrit_parent_trimdata_invgamma_10e5 <- model_run(prior_invgamma_params,'speed',"animal+clutch+dam+sire",4,100000)
  #invgamma2
  Ucrit_parent_trimdata_invgamma2_10e5 <- model_run(prior_invgamma2_params,'speed',"animal+clutch+dam+sire",4,100000)
  #flat
  Ucrit_parent_trimdata_flatprior_10e5 <- model_run(prior_flat_params,'speed',"animal+clutch+dam+sire",4,100000)
  #save them as a list
  Ucrit_parent_trimdata_10e5 <- list(Ucrit_parent_trimdata_priorexpand_10e5,Ucrit_parent_trimdata_priorext_10e5,
                                    Ucrit_parent_trimdata_invgamma_10e5,Ucrit_parent_trimdata_invgamma2_10e5,
                                    Ucrit_parent_trimdata_flatprior_10e5)
  save(Ucrit_parent_trimdata_10e5,file='../model_runs/Ucrit_parent_trimdata_10e5.RData')
```


## Run multiple chains for the best prior for each trait and model:

### SL, no parental effects
Choosing between inverse gamma and parameter-expanded prior. Both mix acceptably well, but the parameter-expanded prior gives a bimodal posterior (one peak at 0, another near where the peak is found with the inverse gamma prior). 

Gelman: "The other way a model can be difficult to fit is if its parameters are only weakly constrained by the posterior. With a small sample size (or, in a hierarchical model, a small number of groups), uncertainty in the hyperparameters can yield a posterior distribution of widely varying curvature, which leads to slowly mixing MCMC." (Stacking for Non-mixing Bayesian Computations, arXiv)

Maybe the bimodal posterior is the result of adding an extra parameter (alpha) from the parameter-expanded prior? In that case, seems okay to just go with the inverse gamma prior.

#### inverse gamma prior

```{r,eval=FALSE}
prior_invgamma2 = list(R = list(V = 1, nu = 0.02),
             G = list(G1 = list(V = 1,nu = .02),
                      G2 = list(V = 1,nu = .02)))

start1=list(R=runif(1),G=c(runif(1),runif(1)))
start2=list(R=runif(1),G=c(runif(1),runif(1)))
start3=list(R=runif(1),G=c(runif(1),runif(1)))

SL_noparent_chain1 <- MCMCglmm(length~1+age,
                              random=~animal+clutch,
                              prior=prior_invgamma2,
                              start=start1,
                              ginverse=list(animal=Ainv),
                              data=phens,nitt=100000,burnin=1000,thin=10,verbose=TRUE)

SL_noparent_chain2 <- MCMCglmm(length~1+age,
                                random=~animal+clutch,
                                prior=prior_invgamma2,
                                start=start2,
                              ginverse=list(animal=Ainv),
                                data=phens,nitt=100000,burnin=1000,thin=10,verbose=TRUE)
SL_noparent_chain3 <- MCMCglmm(length~1+age,
                                random=~animal+clutch,
                                prior=prior_invgamma2,
                                start=start3,
                              ginverse=list(animal=Ainv),
                                data=phens,nitt=100000,burnin=1000,thin=10,verbose=TRUE)

SL_noparent_chains_fulldata <- list(SL_noparent_chain1,SL_noparent_chain2,SL_noparent_chain3)
save(SL_noparent_chains_fulldata,file='../model_runs/SL_noparent_chains_fulldata.RData')
```

#### parameter-expanded prior

```{r, eval=FALSE}
#from tuto_en; with residual edited (nu=0.002 instead of nu=1) to be proper
prior_ext<-list(R=list(V=1,nu=0.002),
                G=list(G1=list(V=1,nu=1,alpha.mu=0,alpha.V=1000),
                       G2=list(V=1,nu=1,alpha.mu=0,alpha.V=1000)))

start1=list(R=runif(1),G=c(runif(1),runif(1)))
start2=list(R=runif(1),G=c(runif(1),runif(1)))
start3=list(R=runif(1),G=c(runif(1),runif(1)))

SL_noparent_expand_chain1 <- MCMCglmm(length~1+age,
                              random=~animal+clutch,
                              prior=prior_ext,
                              start=start1,
                              ginverse=list(animal=Ainv),
                              data=phens,nitt=100000,burnin=1000,thin=10,verbose=TRUE)

SL_noparent_expand_chain2 <- MCMCglmm(length~1+age,
                                random=~animal+clutch,
                                prior=prior_ext,
                                start=start2,
                              ginverse=list(animal=Ainv),
                                data=phens,nitt=100000,burnin=1000,thin=10,verbose=TRUE)
SL_noparent_expand_chain3 <- MCMCglmm(length~1+age,
                                random=~animal+clutch,
                                prior=prior_ext,
                                start=start3,
                              ginverse=list(animal=Ainv),
                                data=phens,nitt=100000,burnin=1000,thin=10,verbose=TRUE)

SL_noparent_expand_chains_trim <- list(SL_noparent_expand_chain1,SL_noparent_expand_chain2,SL_noparent_expand_chain3)
save(SL_noparent_expand_chains_trim,file='../model_runs/SL_noparent_expand_chains_trim.RData')
```

