---
title: "Amphiprion percula Larval Trait Heritability"
author: "E Schlatter"
date: "12/18/2021"
output: html_document
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(MasterBayes) #for pedigree functions
library(MCMCglmm) #for running the model
library(MCMCpack) #for inverse wishart distributions
library(bayesplot) #mcmc_areas plot function
```

```{r}
source("../functions/model_fns.R")
#load("../model_runs/")
```

## Read in and process the data.

Creates the following objects we'll use going forward:

$\textbf{phens}$: a list of each larva and its swimming speed and clutch, plus other covariates to be added. (Each parent also has a row in this list, to be compatible with relatedness matrix, but its swimming speed and covariates are all NA.)

$\textbf{Ainv}$: the inverse of the relatedness matrix, A. In A, $A_{ij}=$probability of identity by descent of an allele in individuals i and j.
```{r data cleanup}
#swim trial data
larvae <- as.data.frame(read.csv('../data/Larvae.csv',header=TRUE)) %>%
  mutate(larvalID=Clutch+.01*Larva) #create unique identifier for each larva

#clutch data
metadata <- as.data.frame(read.csv('../data/metadata.csv',header=TRUE))
colnames(metadata)[1]='Clutch'

#length data
lengths <- as.data.frame(read.csv('../data/Lengths.csv',header=TRUE)) %>%
  mutate(larvalID=Clutch+.01*Larva) %>% #create unique identifier for each larva
  dplyr::select(-Clutch,-Larva) %>%
  rename(Measurer=Initials) %>%
  rename(SL=Average.length) %>%
  mutate(SL=as.numeric(SL))  #this will give "NAs introduced by coercion" -- that's fine

for(i in 1:nrow(larvae)){ #remove SL measurements for dead larvae; they're not accurate
  larvae$SL[i]=ifelse(larvae$Dead.before.preservation.[i]==1,NA,larvae$SL[i])} 

#merge swim trial and length data
larvae <- merge(larvae,lengths,by='larvalID')

larvae <- larvae %>% mutate('Time'=Time.min.+(1/60)*Time.sec.) %>% #add a Time column
  subset(Clutch>4) %>% #first four clutches were practice, and we won't use them in the dataset
  mutate(Clutch=as.factor(Clutch)) %>%
  mutate(larvalID=as.factor(larvalID))

#merge the larvae and metadata data frames, so metadata info is associated with each larva
larvae <- merge(larvae,metadata,by='Clutch') %>%
  dplyr::select(larvalID,everything()) #move larvalID to first column

larvae <- larvae %>% subset(is.na(SL)==FALSE) %>% #remove larvae with no SL measurement
  subset(is.na(Time)==FALSE) %>% #remove larvae with no speed measurement
  mutate(speed=Time-2) #Ucrit is swim time minus 2cm/sec

#make sure all covariates are in the right format
larvae <- larvae %>%
  mutate(R1.ID.=as.factor(R1.ID.)) %>%
  mutate(R2.ID.=as.factor(R2.ID.)) %>%
  mutate(Date=as.Date(Date, "%d-%b-%y")) %>%
  mutate(PostSwitch=as.factor(PostSwitch)) %>%
  mutate(Parental.tank=as.factor(Parental.tank))
```
Trim the dataset to larvae with both maternal and paternal half-sibs. 581 larvae from 12 mothers and 12 fathers.
```{r trim data, eval=FALSE}
mothers_trim <- c(5,6,9,10,13,18,22,30,59,66,70,114) #mothers in the trimmed dataset
fathers_trim <- c(38,96,97,98,103,108,111,115,137,156,165,171) #fathers in the trimmed dataset
  #error in original data entry: father #137 is sometimes written #131
metadata <- subset(metadata,R1.ID.%in%mothers_trim & R2.ID.%in%fathers_trim)
larvae <- subset(larvae,R1.ID.%in%mothers_trim & R2.ID.%in%fathers_trim)
```

Create:

  - pedigree (and full_pedig, including the parents)

  - Ainv

  - phens (animal, speed, length, covariates)

```{r}
#pedigree and Ainv -- we'll need this later for heritability analysis
pedigree <- dplyr::select(larvae,id=larvalID,dam=R1.ID.,sire=R2.ID.)

#pedigree utility functions to insert parents and correctly order, from MasterBayes.
full_pedig <- pedigree %>%
  insertPed() %>%
  orderPed()
Ainv <- MCMCglmm::inverseA(full_pedig)$Ainv

#phenotype data frame
phens <- dplyr::select(larvae,animal=larvalID,speed=speed, length=SL, clutch=Clutch, 
                       dam=R1.ID., sire=R2.ID., tank=Parental.tank, age=Age.at.hatch)
  #add empty rows to phens for parents
parents <- subset(full_pedig,is.na(dam)) %>%
  transmute(animal=id,speed=as.numeric(dam),length=as.numeric(dam),clutch=NA,
            dam=NA,sire=NA,tank=NA,age=1)
phens <- as.data.frame(phens)
phens <- rbind(parents,phens)
```

## Animal model

### Priors
From Hadfield course notes (p.13):
"The inverse gamma is a special case of the inverse Wishart, although it is parametrised using shape and scale, where nu = 2$*$shape and V = scale/shape (or shape = nu/2 and scale = nu$*$V/2 ). MCMCpack provides a density function (dinvgamma) for the inverse gamma distribution."

"The distribution tends to a point mass on V as the degree of belief parameter, nu, goes to infinity." The mean is not define for nu<2.

Several priors to try:
```{r priors, results='hide'}

#parameter-expanded prior
prior_expand_params <- list(V = 1,nu = 1000, alpha.mu=0, alpha.V=1)

#also parameter-expanded, from tuto_en; with residual edited (nu=0.002 instead of nu=1) to be proper
prior_ext_params <- list(V=1,nu=1,alpha.mu=0,alpha.V=1000)

#inverse gamma prior
prior_invgamma_params <- list(V = 1,nu = .002)

#gentler inverse gamma prior
prior_invgamma2_params <- list(V = 1,nu = .02)

#flat prior
prior_flat_params <- list(V = 1e-10,nu = -1) #trusting based on course notes that this is actually flat, because I can't tell what a negative nu means
```


Plot them (when possible):
```{r}
#invgamma
param_list <- prior_invgamma_params
nu <- param_list$nu
V <- param_list$V
#corresponding shape and scale (invgamma) parameters
shape=nu/2
scale=nu*V/2

#invgamma2
param_list.2 <- prior_invgamma2_params
nu.2 <- param_list.2$nu
V.2 <- param_list.2$V
#corresponding shape and scale (invgamma) parameters
shape.2=nu.2/2
scale.2=nu.2*V.2/2

xrange <- seq(0.001,1,.001)

ggplot()+
  geom_line(aes(x=xrange,y=MCMCpack::dinvgamma(xrange,shape,scale),
                color='invgamma'))+
  geom_line(aes(x=xrange,y=MCMCpack::dinvgamma(xrange,shape.2,scale.2),
                color='invgamma2'))+
  labs(x="sigma^2",y="probability",title='prior distributions')
```

## Run models

```{r, eval=FALSE}
##test to make sure it works
param_list=prior_invgamma2_params
model_statement="animal+clutch"
n_vars=2 #not including the residual (units)
n_mcmc_iter=100000


model <- model_run(prior_invgamma2_params,'speed',"animal+clutch",2,1000)
save(model,file='../model_runs/test.RData')
diagn <- model_diagnostics(model)
results <- model_results(model)
```

## Full dataset models

```{r}
### SL, no parental effects

#prior_expand
SL_noparent_fulldata_priorexpand_10e5 <- model_run(prior_expand_params,'length',"animal+clutch",2,100000)

#prior_ext
SL_noparent_fulldata_priorext_10e5 <- model_run(prior_ext_params,'length',"animal+clutch",2,100000)

#invgamma
SL_noparent_fulldata_invgamma_10e5 <- model_run(prior_invgamma_params,'length',"animal+clutch",2,100000)

#invgamma2
SL_noparent_fulldata_invgamma2_10e5 <- model_run(prior_invgamma2_params,'length',"animal+clutch",2,100000)

#flat
SL_noparent_fulldata_flatprior_10e5 <- model_run(prior_flat_params,'length',"animal+clutch",2,100000)

#save them as a list
SL_noparent_fulldata_10e5 <- list(SL_noparent_fulldata_priorexpand_10e5,SL_noparent_fulldata_priorext_10e5,
                                  SL_noparent_fulldata_invgamma_10e5,SL_noparent_fulldata_invgamma2_10e5,
                                  SL_noparent_fulldata_flatprior_10e5)
save(SL_noparent_fulldata_10e5,file='../model_runs/SL_noparent_fulldata_10e5.RData')


### SL, with parental effects

#prior_expand
SL_parent_fulldata_priorexpand_10e5 <- model_run(prior_expand_params,'length',"animal+clutch+dam+sire",4,100000)

#prior_ext
SL_parent_fulldata_priorext_10e5 <- model_run(prior_ext_params,'length',"animal+clutch+dam+sire",4,100000)

#invgamma
SL_parent_fulldata_invgamma_10e5 <- model_run(prior_invgamma_params,'length',"animal+clutch+dam+sire",4,100000)

#invgamma2
SL_parent_fulldata_invgamma2_10e5 <- model_run(prior_invgamma2_params,'length',"animal+clutch+dam+sire",4,100000)

#flat
SL_parent_fulldata_flatprior_10e5 <- model_run(prior_flat_params,'length',"animal+clutch+dam+sire",4,100000)

#save them as a list
SL_parent_fulldata_10e5 <- list(SL_parent_fulldata_priorexpand_10e5,SL_parent_fulldata_priorext_10e5,
                                  SL_parent_fulldata_invgamma_10e5,SL_parent_fulldata_invgamma2_10e5,
                                  SL_parent_fulldata_flatprior_10e5)
save(SL_parent_fulldata_10e5,file='../model_runs/SL_parent_fulldata_10e5.RData')


### Ucrit, no parental effects

#prior_expand
Ucrit_noparent_fulldata_priorexpand_10e5 <- model_run(prior_expand_params,'speed',"animal+clutch",2,100000)

#prior_ext
Ucrit_noparent_fulldata_priorext_10e5 <- model_run(prior_ext_params,'speed',"animal+clutch",2,100000)

#invgamma
Ucrit_noparent_fulldata_invgamma_10e5 <- model_run(prior_invgamma_params,'speed',"animal+clutch",2,100000)

#invgamma2
Ucrit_noparent_fulldata_invgamma2_10e5 <- model_run(prior_invgamma2_params,'speed',"animal+clutch",2,100000)

#flat
Ucrit_noparent_fulldata_flatprior_10e5 <- model_run(prior_flat_params,'speed',"animal+clutch",2,100000)

#save them as a list
Ucrit_noparent_fulldata_10e5 <- list(Ucrit_noparent_fulldata_priorexpand_10e5,Ucrit_noparent_fulldata_priorext_10e5,
                                  Ucrit_noparent_fulldata_invgamma_10e5,Ucrit_noparent_fulldata_invgamma2_10e5,
                                  Ucrit_noparent_fulldata_flatprior_10e5)
save(Ucrit_noparent_fulldata_10e5,file='../model_runs/Ucrit_noparent_fulldata_10e5.RData')


### Ucrit, with parental effects

#prior_expand
Ucrit_parent_fulldata_priorexpand_10e5 <- model_run(prior_expand_params,'speed',"animal+clutch+dam+sire",4,100000)

#prior_ext
Ucrit_parent_fulldata_priorext_10e5 <- model_run(prior_ext_params,'speed',"animal+clutch+dam+sire",4,100000)

#invgamma
Ucrit_parent_fulldata_invgamma_10e5 <- model_run(prior_invgamma_params,'speed',"animal+clutch+dam+sire",4,100000)

#invgamma2
Ucrit_parent_fulldata_invgamma2_10e5 <- model_run(prior_invgamma2_params,'speed',"animal+clutch+dam+sire",4,100000)

#flat
Ucrit_parent_fulldata_flatprior_10e5 <- model_run(prior_flat_params,'speed',"animal+clutch+dam+sire",4,100000)

#save them as a list
Ucrit_parent_fulldata_10e5 <- list(Ucrit_parent_fulldata_priorexpand_10e5,Ucrit_parent_fulldata_priorext_10e5,
                                  Ucrit_parent_fulldata_invgamma_10e5,Ucrit_parent_fulldata_invgamma2_10e5,
                                  Ucrit_parent_fulldata_flatprior_10e5)
save(Ucrit_parent_fulldata_10e5,file='../model_runs/Ucrit_parent_fulldata_10e5.RData')

```


