---
title: "Notes on priors"
author: "E Schlatter"
date: "1/14/2022"
output: html_document
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(extraDistr) #has inverse gamma
library(MCMCpack) #for inverse gamma and inverse wishart distributions
```

```{r}
source("../functions/model_fns.R")
source("priors.R")
```

## Priors
From Hadfield course notes (p.13):
"The inverse gamma is a special case of the inverse Wishart, although it is parametrised using shape and scale, where nu = 2$*$shape and V = scale/shape (or shape = nu/2 and scale = nu$*$V/2 ). MCMCpack provides a density function (dinvgamma) for the inverse gamma distribution."

"The distribution tends to a point mass on V as the degree of belief parameter, nu, goes to infinity." The mean is not define for nu<2.

Several priors to try:
```{r priors, results='hide'}

#parameter-expanded prior
prior_expand_params <- list(V = 1,nu = 1000, alpha.mu=0, alpha.V=1)

#also parameter-expanded, from tuto_en; with residual edited (nu=0.002 instead of nu=1) to be proper
prior_ext_params <- list(V=1,nu=1,alpha.mu=0,alpha.V=1000)

#inverse gamma prior
prior_invgamma_params <- list(V = 1,nu = .002)

#gentler inverse gamma prior
prior_invgamma2_params <- list(V = 1,nu = .02)

#flat prior
prior_flat_params <- list(V = 1e-10,nu = -1) #trusting based on course notes that this is actually flat, because I can't tell what a negative nu means
```

Plot them (when possible):
```{r}
#invgamma
param_list <- prior_invgamma_params
nu <- param_list$nu
V <- param_list$V
#corresponding shape and scale (invgamma) parameters
shape=nu/2
scale=nu*V/2

#invgamma2
param_list.2 <- prior_invgamma2_params
nu.2 <- param_list.2$nu
V.2 <- param_list.2$V
#corresponding shape and scale (invgamma) parameters
shape.2=nu.2/2
scale.2=nu.2*V.2/2

xrange <- seq(0.001,.02,.001)

ggplot()+
  geom_line(aes(x=xrange,y=MCMCpack::dinvgamma(xrange,shape,scale),
                color='invgamma'))+
  geom_line(aes(x=xrange,y=MCMCpack::dinvgamma(xrange,shape.2,scale.2),
                color='invgamma2'))+
  labs(x="sigma^2",y="probability",title='prior distributions')
```

##scratch: inverse Wishart

```{r}
#prior
prior_invgamma_params <- list(V = 1,nu = 0.000001)
param_list <- prior_invgamma_params
nu <- param_list$nu
V <- param_list$V
#corresponding shape and scale (invgamma) parameters
shape=nu/2
scale=nu*V/2

xrange <- seq(0.0001,1,.0001)

#calculate inverse gamma
y_invgamma=MCMCpack::dinvgamma(xrange,shape,scale)
ymax=max(y_invgamma)

#calculate inverse wishart
distys <- data.frame(x=xrange,y=NA,dist='invgamma')
for(i in 1:nrow(distys)){distys$y[i]=diwish(distys$x[i],V,nu)}
distys_scaled <- ymax*distys$y/max(distys$y) #scale it to have the same maximum as the inverse gamma version


ggplot()+
  #geom_line(aes(x=xrange, y=extraDistr::dinvgamma(xrange,alpha=shape,beta=scale)))+
    # #use extraDistr invgamma function
  geom_line(aes(x=xrange,y=y_invgamma))#+
    #equivalently, MCMCpack invgamma function. 
    #These two also match the plots in Hadfield course notes (p.12)

#  geom_line(aes(x=xrange,y=distys_scaled),color='blue')
    # #using diwish from MCMCpack -- NOT equivalent. much steeper.
    # #(and peak is super high ~300 if you don't scale it)
```

